---
title: "MPS Project -- Exploratory Analysis"
author: "Sung-Woo Ahn Wanwen Gu Yvonne Liu Xinyue Chen"
date: "March 8, 2019"
output: 
  html_document:
    highlight: pygments
    theme: cosmo
---

<style type="text/css">

body{ /* Normal  */
      font-size: 18px;
  }

</style>



## Import dataset
```{r warning=FALSE, message=FALSE, echo=FALSE}
library(knitr)
library(readr)
library(data.table)
library(bit64)
library(Rmpfr)
library(tidyverse)
library(plyr)
library(ggmosaic)
library(ggmap)
library(plotly)
library(kableExtra)
library(readxl)

```
```{r echo=FALSE}
options(kableExtra.latex.load_packages = FALSE)
knit_hooks$set(document = function(x) {
  sub('\\usepackage[]{color}', '\\usepackage[]{xcolor}', x, fixed = TRUE)
})

```

```{r}
train <- fread("train.csv", header = T)
setDT(train[,-1])
labelsdat <- fread("labelsdat.csv", header = T)
labeltrain <- fread("labeltrain.csv", header = T)
```

## Exploratory Analysis

### Missing Values

#### Location Data
```{r}
dim(train[is.na(longitude) | is.na(latitude) | longitude == "" | latitude == ""])

inva_locations<-sum(train$longitude == 1 | train$longitude ==0 | train$latitude ==1 | train$latitude ==0)/length(train$longitude)
```
No missing data, but 58% of the location data is still invalid.

#### Phone Brand and Model Data
```{r}
dim(train[train$phone_brand==""])
dim(train[train$phone_brand==""])[1]/dim(train)[1]*100

dim(train[train$device_model==""])
dim(train[train$device_model==""])[1]/dim(train)[1]*100
```
Almost 100% of phonebrands and phone models are missing

#### Is_Installed and Is_Active Data
```{r}
dim(train[is.na(is_installed)])
dim(train[is.na(is_installed)])[1]/dim(train)[1]*100

dim(train[is.na(is_active)])
dim(train[is.na(is_active)])[1]/dim(train)[1]*100
```
5.11% missing in is_installed and is_active variables

```{r}
dim(train[is.na(app_id) | app_id == ""])
dim(train[is.na(app_id)])[1]/dim(train)[1]*100
```
5.11% missing in app_id

#### The rest: no missing values
```{r}
dim(train[is.na(timestamp) | timestamp == ""])
dim(train[is.na(event_id) | event_id == ""])
dim(train[is.na(device_id) | device_id == ""])
dim(train[is.na(gender) | gender == ""])
dim(train[is.na(age) | age == ""])
dim(train[is.na(group) | group == ""])
```

#### Relationship of Gender and Age with Event_id and Device_id
```{r fig.width=13}
train2 <- train
require(data.table)
setDT(train2)[, count := uniqueN(event_id), by = device_id]
count_event = train2[, list(device_id, group,count)]
count_event2 = unique(count_event)  # length 23309

grp <- c('F23-','F24-26','F27-28','F29-32','F33-42','F43+','M22-','M23-26','M27-28','M29-31','M32-38','M39+')
len <- length(grp)
grp_num <- rep(0, len)
grp_sum <- rep(0, len)
grp_mean <- rep(0, len)
grp_median <- rep(0, len)
grp_max <- rep(0, len)
grp_min <- rep(0, len)

for (i in 1:len){
  thisgroup <- count_event2[group==grp[i]]
  grp_num[i] <- dim(thisgroup)[1]
  grp_sum[i] <- sum(thisgroup$count)
  grp_mean[i] <- mean(thisgroup$count)
  grp_median[i] <- median(thisgroup$count)
  grp_max[i] <- max(thisgroup$count)
  grp_min[i] <- min(thisgroup$count)
}

grp_event <- data.frame('group'=grp,'number'=grp_num,'sum'=grp_sum,'mean'=grp_mean,'median'=grp_median,
                        'max'=grp_max,'min'=grp_min)
```

```{r}
#the same as dataset grp_event
test=ddply(count_event2,"group",summarise,n=length(device_id),sum=sum(count),mean=mean(count),median=median(count),max=max(count),min=min(count))
```

```{r results='asis'}
grp_event %>%
  kable(escape = F, caption = 'Events') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```

```{r}
barplot(grp_sum,names.arg=grp, las=2, xlab = 'Group', main = 'Event Distribution by Group')
barplot(grp_num,names.arg=grp, las=2, xlab = 'Group', main = 'Unique Device ID by Group')
barplot(grp_mean,names.arg=grp, las=2, xlab = 'Group', main = 'Mean Event Count by Group')
```

### Univariate Exploratory Analysis
#### Gender Distribution
```{r}
genderbyevent <- train %>%
  group_by(gender) %>%
  dplyr::summarise(n= n()) %>%
  mutate(percent = n/sum(n))

ggplot(data=genderbyevent,aes(x=gender, y = percent, fill=gender))+
  geom_bar(stat="identity")+
  ggtitle("Gender Distribution by Event Count")
```
```{r}
genderbydevice <- train %>%
  select(device_id, gender) %>%
  distinct(device_id, .keep_all = T) %>%
  group_by(gender) %>%
  dplyr::summarise(n= n()) %>%
  mutate(percent = n/sum(n))

ggplot(data=genderbydevice,aes(x=gender, y = percent, fill=gender))+
  geom_bar(stat="identity")+
  ggtitle("Gender Distribution by Device Count")
```
There are more males than females in the data.  Can also see that male counterparts use their devices more frequently since the distribution is more skewed in the by event graph.


#### Age-Gender Distribution
```{r warning=FALSE}
ggplot(data = train, aes(x = age,fill = gender, color = gender))+
  geom_histogram(stat = "count",alpha = .5 ,position = "identity")+
  ggtitle("Age-Gender Distribution by Event Count")

agebydevice <- train %>%
  select(device_id, gender, age) %>%
  distinct(device_id, .keep_all = T)

ggplot(data = agebydevice, aes(x = age, fill = gender, color = gender)) +
  geom_histogram(stat = "count",alpha = .5 ,position = "identity")+
  ggtitle("Age-Gender Distribution by Device Count")
```
Both charts show a right skewed distribution. Would this be a problem for prdictive modeling?


#### Group Distribution
```{r}
ggplot(data=train,aes(x=group,fill=gender))+
  geom_bar(stat="count")+
  ggtitle("Group Distribution by Event")

groupbydevice <- train %>%
  select(device_id, gender, group) %>%
  distinct(device_id, .keep_all = T)

ggplot(data=groupbydevice,aes(x=group,fill=gender))+
  geom_bar(stat="count")+
  ggtitle("Group Distribution by Device")
```

#### Is_Active Distribution
```{r}
isactiveperc <- train %>%
  group_by(is_active) %>%
  dplyr::summarise(n= n()) %>%
  mutate(percent = n/sum(n))

ggplot(data=isactiveperc,aes(x=factor(is_active), y = percent))+
  geom_bar(stat="identity", fill="steelblue")+
  ggtitle("Is_Active Count")+
  labs(x = "is_active", y = "Count")
```


#### Phone Brand Distribution
```{r}
phone_count <- train %>%
  group_by(phone_brand) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```
```{r}
phone_count2 <- train %>%
  group_by(phone_brand) %>%
  distinct(device_id) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```
```{r results='asis', echo=FALSE}
kable(phone_count2[2:21,], caption = 'Phone Devices Count') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive", "condensed"), full_width = F)
```

```{r eval = FALSE, echo=FALSE}
# Find phonebrand count
phonebrandsorted <- train %>%
  select(phone_brand) %>%
  group_by(phone_brand) %>%
  dplyr::summarize(count = n()) %>%
  filter(phone_brand != "") %>%
  arrange(desc(count))
```
```{r fig.width=11}
ggplot(data = phone_count2[2:21,], aes(x = reorder(phone_brand, -count), y = count, fill=count)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 18)) +
  labs(x = 'Phone Brands', y = 'Count', title = 'Phone Brand Distribution')
```

#### App category counts
```{r fig.width=17, fig.height=8}

response <- train[, list(device_id, gender, age, group, app_id)]

uniquelabels <- labelsdat %>%
  arrange(app_id) %>%
  group_by(app_id, category) %>%
  distinct(category)

labeltrain <- merge(x = response, y = uniquelabels, by = 'app_id', all.x = T, allow.cartesian = T)

# Find label count
labelcount <- uniquelabels %>%
  group_by(category) %>%
  dplyr::summarise(count = n()) %>%
  arrange(desc(count))

ggplot(data = labelcount, aes(x = reorder(category,-count), y=count, fill = count)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 15),
        axis.text.y = element_text(size = 15),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25)) +
  labs(x = 'App Categories', y = 'Count', title = 'App Category Distribution')
```


#### Map of the location data
```{r fig.width=11, fig.height=7, fig.align='center', message=FALSE, warning=FALSE, eval=FALSE}
locations <- train %>%
  select(longitude, latitude) %>%
  filter(longitude != 1 & longitude != 0 & latitude !=1 & latitude != 0)
box <- make_bbox(lon = locations$longitude, lat = locations$latitude, f = .1)

map <- get_map(location = box, maptype = 'terrain', source = 'google')

p <- ggmap(map)
p + geom_point(data = locations, mapping = aes(x = longitude, y = latitude), color = "red")
```

Most events occurred in Asia

### Bivariate Exploratory Analysis
#### Relationship of Phonebrand with Gender & Age
```{r echo = FALSE}
# Function to count the number of different brands each group category uses
phone_count=function(group) {
  lenova<-dim(train[which(train$group==group& train$phone_brand=='lenova')])[1]
  samsung<-dim(train[which(train$group==group & train$phone_brand=='samsung')])[1]
  ctyon<-dim(train[which(train$group==group & train$phone_brand=='ctyon')])[1]
  coolpad<-dim(train[which(train$group==group & train$phone_brand=='coolpad')])[1]
  vivo<-dim(train[which(train$group==group & train$phone_brand=='vivo')])[1]
  huawei<-dim(train[which(train$group==group & train$phone_brand=='huawei')])[1]
  gionee<-dim(train[which(train$group==group & train$phone_brand=='gionee')])[1]
  le<-dim(train[which(train$group==group & train$phone_brand=='le')])[1]
  meizu<-dim(train[which(train$group==group & train$phone_brand=='meizu')])[1]
  mi<-dim(train[which(train$group==group & train$phone_brand=='mi')])[1]
  meitu<-dim(train[which(train$group==group & train$phone_brand=='meitu')])[1]
  smartisan<-dim(train[which(train$group==group & train$phone_brand=='smartisan')])[1]
  yitong<-dim(train[which(train$group==group & train$phone_brand=='yitong')])[1]
  zte<-dim(train[which(train$group==group & train$phone_brand=='zte')])[1]
  koobee<-dim(train[which(train$group==group & train$phone_brand=='koobee')])[1]
  doov<-dim(train[which(train$group==group & train$phone_brand=='doov')])[1]
  nubia<-dim(train[which(train$group==group & train$phone_brand=='nubia')])[1]
  na<-dim(train[which(train$group==group & is.na(train$phone_brand))])[1]
  group_phone_count<-data.frame('phone_brand'=c('lenova','samsung', 'ctyon', 'coolpad', 'vivo', 'huawei', 'gionee', 'le','meizu','smartisan','yitong', 'zte', 'koobee', 'doov','nubia', 'na'), 'count'=c(lenova, samsung, ctyon, coolpad, vivo, huawei, gionee, le, meizu, smartisan, yitong, zte, koobee, doov, nubia, na))
  return(group_phone_count)
}

```
```{r}
group_phone_count <- train %>%
  group_by(group, phone_brand) %>%
  distinct(device_id) %>%
  dplyr::summarise(count = n()) %>%
  arrange(desc(count), .by_group = TRUE)
```

```{r results='asis', echo=FALSE}
kable(group_phone_count, caption = 'Phone Brands By Group') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F) %>%
  group_rows("F23-", 1, 6) %>%
  group_rows("F24-26", 7, 11) %>%
  group_rows("F27-28", 12, 15) %>%
  group_rows("F29-32", 16, 21) %>%
  group_rows("F33-42", 22, 27) %>%
  group_rows("F43+", 28, 32) %>%
  group_rows("M22-", 33, 38) %>%
  group_rows("M23-26", 39, 48) %>%
  group_rows("M27-28", 49, 56) %>%
  group_rows("M29-31", 57, 67) %>%
  group_rows("M32-38", 68, 77) %>%
  group_rows("M39+", 78, 85)
```

```{r echo=FALSE}
phonebrandmf <- train %>%
  mutate(mf = ifelse(grepl("M", train$group) == TRUE, "Male", "Female")) %>%
  filter(phone_brand != "") %>%
  select(group, mf, phone_brand)
```




#### Relationship of App Usage (event_id not unique) with Gender & Age
```{r}
# Count number of app uses each group category has
m39 <- dim(train[which(train$group == 'M39+')])[1]
m32.38<- dim(train[which(train$group == 'M32-38')])[1]
m29.31<- dim(train[which(train$group == 'M29-31')])[1]
m27.28 <- dim(train[which(train$group == 'M27-28')])[1]
m23.26 <- dim(train[which(train$group == 'M23-26')])[1]
m22 <- dim(train[which(train$group == 'M22-')])[1]
f43 <- dim(train[which(train$group == 'F43+')])[1]
f33.42 <- dim(train[which(train$group == 'F33-42')])[1]
f29.32 <- dim(train[which(train$group == 'F29-32')])[1]
f27.28 <- dim(train[which(train$group == 'F27-28')])[1]
f24.26 <- dim(train[which(train$group == 'F24-26')])[1]
f23 <- dim(train[which(train$group == 'F23-')])[1]

groupcts <- data.frame('category' = c('M39+', 'M32-38', 'M29-31', 'M27-28', 'M23-26', 'M22-', 'F43+', 'F33-42', 'F29-32', 'F27-28', 'F24-26', 'F23-'), 'counts' = c(m39, m32.38, m29.31, m27.28, m23.26, m22, f43, f33.42, f29.32, f27.28, f24.26, f23))


# Simpler with dplyr
groupcts2 <- train %>%
  group_by(group) %>%
  dplyr::summarise(count = n()) %>%
  mutate(proportion = 100*count/sum(count))
```

```{r results='asis'}
groupcts2 %>%
  mutate(
    count = cell_spec(count, color = ifelse(count == max(count[1:6]) | count == max(count[7:12]), "blue", "black")),
    proportion = cell_spec(proportion, color = ifelse(proportion == max(proportion[1:6]) | proportion == max(proportion[7:12]), "blue", "black"))
  ) %>%
kable(escape = F, caption = 'App Usage Count by Group') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```

Visualization App Usage
```{r}
# Dplyr way
groupgendercts <- train %>%
  select(gender, age) %>%
  mutate(agecat = case_when(age <= 23 ~ "23-",
                            age > 23 & age <= 26 ~ "24-26",
                            age > 26 & age <= 28 ~ "27-28",
                            age > 28 & age <= 32 ~ "29-32",
                            age > 32 & age <= 42 ~ "33-42",
                            age > 42 ~ "43+"))

# data.table way
groupgendercts <- train[,list(gender, age, agecat = case_when(age <= 23 ~ "23-",
                            age > 23 & age <= 26 ~ "24-26",
                            age > 26 & age <= 28 ~ "27-28",
                            age > 28 & age <= 32 ~ "29-32",
                            age > 32 & age <= 42 ~ "33-42",
                            age > 42 ~ "43+"))]


ggplot(data = groupgendercts) +
  geom_mosaic(aes(x = product(gender, agecat), fill = gender)) +
  labs(x = "Age", y = "Gender", title = "App Usage by Age and Gender")
```

