---
title: "MPS Project -- Data Cleaning and Merging"
author: "Sung-Woo Ahn Wanwen Gu Yvonne Liu Xinyue Chen"
date: "March 8, 2019"
output:
  html_document:
    highlight: pygments
    theme: cosmo
---

<style type="text/css">

body{ /* Normal  */
      font-size: 18px;
  }

</style>

In this document, go over how we cleaned and merged the 6 datasets:

- gender_age_train.csv
- events.csv
- app_events.csv
- phone_brand_device_model.csv
- label_categories.csv
- app_labels.csv

Read in libraries
```{r warning=FALSE, message=FALSE}
library(knitr)
library(readr)
library(readxl)
library(data.table)
library(bit64)
library(Rmpfr)
library(tidyverse)
library(plyr)
library(kableExtra)
```
```{r echo=FALSE}
options(kableExtra.latex.load_packages = FALSE)
knit_hooks$set(document = function(x) {
  sub('\\usepackage[]{color}', '\\usepackage[]{xcolor}', x, fixed = TRUE)
})

```

## Data Import

```{r message=FALSE}
# Plain read.csv can take a while. fread from data.table library is much more efficient
events <- fread(file = 'events.csv', header = T, integer64 = "character")
app_events <- fread('app_events.csv', header = T, integer64 = "character")
app_labels <- fread('app_labels.csv', header = T, integer64 = "character")
labelcategory <- fread('label_standardized categories.csv', header = T, integer64 = "character")
gender_age <- fread('gender_age_train.csv', header = T, integer64 = "character")


# Use read_csv here because of the Chinese characters issue
phonebrand <- read_csv('phone_brand_device_model.csv', locale = locale(encoding = 'UTF-8'))
# Manually convert device_id to character type
phonebrand$device_id <- as.character(phonebrand$device_id)
```

- The id columns like app_id and device_id are converted to character type for efficiency
- There is an issue with R not recognizing Chinese characters in the phone_brand_device_model file. Change the locale setting in read_csv to recognize Chinese characters. Below, we also translated all columns with Chinese to English for easier analysis.

## Data Cleaning

### Translate Chinese characters in the phone brand and model columns

Created two excel files with translation keys; one for phone brand, and one for device model
```{r}
# Read in the translation dataset
chin_eng_brand <- read_excel("phone_brand-translate.xlsx", sheet = "Sheet1")
chin_eng_model <- read_excel("device_model translate.xlsx", sheet="Sheet1")
```
```{r results='asis'}
kable(chin_eng_brand[1:10,], caption = 'Preview of Phone Brand Translation Key') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
kable(chin_eng_model[1:10,], caption = 'Preview of Phone Model Translation Key') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```

Use a for loop: if the Chinese phone brand or model name matches a name on our translation key, replace with its English equivalent
```{r}
# Replace each brand name with an english one with for loop
phone <- phonebrand$phone_brand
for (i in 1:121) {
  phone[phone==chin_eng_brand$`original phone brand`[i]] = chin_eng_brand$`new phone brand`[i]
}
phonebrand$phone_brand <- phone

# Replace each model name with an english one with for loop
model<-phonebrand$device_model
for (i in 1:178){
  model[model==chin_eng_model$`original model name`[i]]=chin_eng_model$`new model name`[i]
}
phonebrand$device_model <- model
```

Now check for any rows where Chinese characters are still left. If there are no alphabets, then basically assume its still Chinese character
```{r}
# Check to make sure that all Chinese characters have been replaced
leftoversbr <- phonebrand[which(grepl("[[:alnum:]]", phonebrand$phone_brand) == FALSE), ]
leftoversbr

leftoversmod <- phonebrand[which(grepl("[[:alnum:]]", phonebrand$device_model) == FALSE), ]
leftoversmod
```
Everything was translated.


## Merging

Before merging, look at the properties of each of the datasets. Determine which columns are unique and which have duplicates.

**Main Points:**

- gender_age: Dataset with a list of device_id and their response, gender_age. **Unique** device_id.
- events: Dataset with a list of all events. **Unique** event_id and nonunique device_id.
- App_events: Dataset with a list of apps used at each event. App_id and event_id not unique.
- phonebrand: Dataset with a list of device_id and their phone brand and model. Device_id not unique, although many of the duplicates have exactly same information.
- app_labels: Dataset with a list of all apps and their label_id. App_id and label_id not unique.
- Labelcategory: Dataset with a list of label_id and their category. **Unique** label_id, category and subcategory not unique.

### events dataset
```{r}
# Any duplicate event_id? NO
dim(events[which(duplicated(events$event_id) == T),])
# Any duplicated device_id? YES
dim(events[which(duplicated(events$device_id) == T),])
```
```{r results='asis'}
# An of the example duplicates
kable(events[device_id==events[[81, 2]]][1:10,], caption = 'One device_id can have many Events') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
This dataset has a list of unique events, but multiple events could have same device_ids. The event_id is unique index.

### app_events dataset
```{r}
# Any duplicate event_id? YES
dim(app_events[which(duplicated(app_events$event_id) == T),])
# Any duplicate app_id? YES
dim(app_events[which(duplicated(app_events$app_id) == T),])
```
```{r results='asis'}
# An example of the duplicates
kable(app_events[event_id == 568,][1:10], caption = 'Multiple Apps Used in One Event') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
kable(app_events[app_id == '6284164581582112235',][1:10], caption = 'Same Apps are used in Multiple Events') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
For every event_id, there are multiple app_ids associated. Same app_ids can be associated with different event_ids. This dataset has no unique index.

### app_labels dataset
```{r}
# Any duplicate app_id? YES
dim(app_labels[which(duplicated(app_labels$app_id) == T),])
# Any duplicated label_id? YES
dim(app_labels[which(duplicated(app_labels$label_id) == T),])
# An of the example duplicates
```
```{r results='asis'}
kable(app_labels[app_id==app_labels[[4, 1]],][1:10], caption = 'Same App is Under Multiple Categories') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)

kable(app_labels[label_id==app_labels[[5, 2]],][1:10], caption = 'Multiple Apps Are Under Same Category') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
App_ids can be categorized under multiple label_ids, and the same label_ids are used multiple times for different app_ids. No unique index.

### label_category dataset
```{r}
# Any duplicate label_id? NO
dim(labelcategory[which(duplicated(labelcategory$label_id) == T),])
# Any duplicated category? YES
dim(labelcategory[which(duplicated(labelcategory$`sub-category`) == T),])
```
```{r results='asis'}
# An of the example duplicates
kable(labelcategory[`sub-category`=="Medical"], caption = 'Some Label_id have same sub-category') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)

kable(labelcategory[category=="game"][1:10], caption = 'Many Label_id have same category') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
Label_id is unique. Sub-categories and categries are not unique, so multiple label_ids can have the same sub-category or categories.

### Phonebrand dataset
```{r}
# Any duplicates in device_id? YES
dim(phonebrand[which(duplicated(phonebrand$device_id) == T),])
```
```{r results='asis'}
# Example duplicates
kable(phonebrand[phonebrand$device_id==phonebrand[[22626,1]],], caption = 'Some Device_id are repeated with exact same information') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)

kable(phonebrand[which(phonebrand$device_id==phonebrand[[34351,1]]),], caption = 'Some Device_id are repeated with exact same information') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
Looking at the examples, looks like the duplicates are the exact same individuals. Safe to delete?
```{r}
# Find device_ids that are repeated AND have the same device models
phonebrand2 <- phonebrand %>%
  group_by(device_id, device_model) %>%
  dplyr::summarise(phonebrand.count = n()) %>%
  filter(phonebrand.count > 1) %>%
  arrange(device_id)
dim(phonebrand2)
```
523 device_ids have another row with the same device model
```{r}
# Find device ids with repeating rows
phonebrand3 <- phonebrand %>%
  group_by(device_id) %>%
  dplyr::summarise(device_id.count = n()) %>%
  filter(device_id.count > 1)
dim(phonebrand3)
```
530 device_ids have another row, but they don't have to have the same device model. This means that there are 7 device_ids with 2 different phones brands.
```{r}
twophones <- setdiff(phonebrand3$device_id, phonebrand2$device_id)
```

```{r}
phonebrand[which(phonebrand$device_id %in% twophones),] %>%
  arrange(device_id)
```
In the cases where the device_id have two different device models, should not delete the rows.

### Gender_age dataset
```{r}
# Any duplicate device_id? NO
dim(gender_age[which(duplicated(gender_age$device_id) == T),])
```
```{r results='asis'}
kable(arrange(gender_age, device_id)[1:15,], caption = 'Preview of Gender_Age Dataset') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
Gender_age has a unique device_id column. Every individual has a gender/age class.


### Merge Operations

First delete the duplicate rows in phonebrand dataset, while at the same time keeping the device_ids that have two different phone models. The unique() function will do this for us since it only deletes rows with exact same information.
```{r}
uniquephones <- unique(phonebrand)
```

First left join phonebrand with gender_age by device_id
```{r results='asis'}
kable(gender_age[1:6, ], caption = 'Gender_Age') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F) %>%
  column_spec(1, width = "15em")
kable(phonebrand[1:6,], caption = 'Phonebrand') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
```{r}
# Convert uniquephones to a data.table to prevent any memory issues
merged1 <- merge(x = gender_age, y = setDT(uniquephones), by = 'device_id', all.x = T)
```
The result
```{r results='asis'}
kable(merged1[1:6,], caption = 'Gender_Age + Phonebrand') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F) %>%
  column_spec(1, width = "15em")
```


Second Merge: Left Join the previous dataset with Events by device_id
```{r results='asis'}
kable(events[1:6,], caption = 'Events') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F) %>%
  column_spec(1, width = "15em")
```
```{r}
merged2 <- merge(x = merged1, y = events, by = 'device_id', all.x = T)
```

<mark> Before moving on to the next merge, remove observations where every column except the response columns, $gender$, $age$, and $group$, is missing. </mark>

```{r}
merged2.noNA <- merged2[is.na(merged2$phone_brand)==FALSE | is.na(merged2$event_id)==FALSE]
```
The result
```{r results='asis'}
kable(merged2.noNA[1:6,], caption = 'Previous + Events') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F) %>%
  column_spec(1, width = "12em") %>%
  column_spec(6, width = "6em")
```


The third merge. Left join the previous dataset with App_Events by event_id.
```{r results='asis'}
kable(app_events[1:6,], caption = 'App_Events') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F) %>%
  column_spec(1, width = "15em")
```
Call this third merge the train dataset.
```{r}
train <- merge(x = merged2.noNA, y = app_events, by = 'event_id', all.x = T)
```
The result
```{r results='asis'}
kable(train[1:6,], caption = "Previous + App_Events") %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```

Save the final train dataset
```{r}
fwrite(train, "train.csv")
```

There are still 2 datasets left pertaining to app descriptions. Merged separately to keep the train dataset from getting very large.
```{r results='asis'}
kable(app_labels[1:6,], caption = 'App_Labels') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)

kable(labelcategory[1:6], caption = 'Label_Category') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
Left Join App_labels with Label_category by label_id. Call this labelsdat.
```{r}
# Merge the app_label with label_category
labelsdat <- merge(x = app_labels, y = labelcategory, by = 'label_id', all.x = T)
```
```{r results='asis'}
kable(labelsdat[1:10,], caption = 'App_labels + Label_category') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
Try merging a few columns from the train dataset with the labelsdat dataset.
```{r}
response <- train[, list(device_id, event_id, app_id)]
```
```{r results='asis'}
kable(response[1:10,], caption = 'Subset of Train Dataset') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
Left join subset of train with Labelsdat by App_id. Call this labeltrain
```{r}
labeltrain <- merge(x = response, y = labelsdat, by = 'app_id', all.x = T, allow.cartesian = T)
labeltrain <- arrange(labeltrain, event_id)
```

The result
```{r results='asis'}
kable(labeltrain[1:10,], caption = 'Subset of Train + Labelsdat') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
Save dataset
```{r eval=FALSE}
fwrite(labeltrain, "labeltrain.csv")
fwrite(labelsdat, "labelsdat.csv")
```
```{r echo=FALSE, eval=FALSE}
response <- train[, list(device_id, gender, age, group, app_id)]
```
```{r results='asis', echo=FALSE, eval=FALSE}
kable(response[1:10,], caption = 'Response') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
```{r echo=FALSE, eval=FALSE}
uniquelabels <- labelsdat %>%
  arrange(app_id) %>%
  group_by(app_id, category) %>%
  distinct(category)

```
```{r results='asis', echo=FALSE, eval=FALSE}
kable(uniquelabels[1:10,], caption = 'UniqueLabels') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```
```{r echo=FALSE, eval=FALSE}
labeltrain <- merge(x = response, y = uniquelabels, by = 'app_id', all.x = T, allow.cartesian = T)
```
```{r results='asis', echo=FALSE, eval=FALSE}
kable(labeltrain[1:10,], caption = 'Training+AppCategory') %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F)
```

### Merge Summary

```{r}
# Observation count of first three merges
merge.dims <- data.frame("Table" = c("Gender_Age", "Previous + Phonebrand", "Previous + Events", "Delete Unneeded Rows", "Previous + AppEvents"), "Join Type" = c("Starting Point", "Left Join", "Left Join", "", "Left Join"), "Merge Key" = c("Device ID", "Device ID", "Device ID", "", "Event ID"), "Uniqueness" = c("Unique", "Unique", "Not Unique", "", "Not Unique"), "Obs Count" = c(dim(gender_age)[1], dim(merged1)[1], dim(merged2)[1], dim(merged2.noNA)[1], dim(train)[1]))

# Observation count of the last two merges
merge.dims2 <- data.frame("Table" = c("App_labels", "Previous + Label_category", "Previous + Subset of Train"), "Join Type" = c("Starting Point", "Left Join", "Left Join"), "Merge Key" = c("Label ID", "Label ID", "App ID"), "Uniqueness" = c("Not Unique", "Unique", "Not Unique"), "Obs Count" = c(dim(app_labels)[1], dim(labelsdat)[1], dim(labeltrain)[1]))
```
```{r results='asis'}
kable(merge.dims, caption = "Change in Length of Table for train dataset") %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F) %>%
  column_spec(2, width = "12em")
kable(merge.dims2, caption = "Change in Length of Table for Labeltrain") %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F) %>%
  column_spec(2, width = "12em")
```

```{r}
# Device id count
merge.idcount <- data.frame("Table" = c("Gender_Age", "Previous + Phonebrand", "Previous + Events", "Delete Unneeded Rows", "Previous + AppEvents"), "Join Type" = c("Starting Point", "Left Join", "Left Join", "", "Left Join"), "Merge Key" = c("Device ID", "Device ID", "Device ID", "", "Event ID"), "Uniqueness" = c("Unique", "Unique", "Not Unique", "", "Not Unique"), "Device ID Count" = c(length(unique(gender_age$device_id)), length(unique(merged1$device_id)), length(unique(merged2$device_id)), length(unique(merged2.noNA$device_id)), length(unique(train$device_id))))
```
```{r echo=FALSE, eval=FALSE}
merge.idcount2 <- data.frame("Table" = c("App_labels", "Previous + Label_category", "Previous + Subset of Train"), "Join Type" = c("Starting Point", "Left Join", "Left Join"), "Merge Key" = c("Label ID", "Label ID", "App ID"), "Uniqueness" = c("Not Unique", "Unique", "Not Unique"), "Device ID Count" = c(length(unique(app_labels$app_id)), length(unique(labelsdat$app_id)), length(unique(labeltrain$app_id))))
```
```{r}
merge.idcount3 <- data.frame("Table" = c("App_labels", "Previous + Label_category", "Previous + Subset of Train"), "Join Type" = c("Starting Point", "Left Join", "Left Join"), "Merge Key" = c("Label ID", "Label ID", "App ID"), "Uniqueness" = c("Not Unique", "Unique", "Not Unique"), "Device ID Count" = c("NA", "NA", length(unique(labeltrain$device_id))))
```
```{r results='asis'}
kable(merge.idcount, caption = "Change in Device ID Count for train dataset") %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F) %>%
  column_spec(2, width = "9em")
kable(merge.idcount3, caption = "Device ID Count for labeltrain dataset") %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "responsive"), full_width = F) %>%
  column_spec(2, width = "9em")
    ```
